---
title: "IV v.1.2"
author: "Kính"
output:
    html_document:
        toc: true
        toc_float: true
        code_folding: hide
always_allow_html: true
---

```{r setup, include=FALSE}
source("~/Documents/libs.r")
```

## Skimming the data

Number of obs before and after remove missing in either death status, survival
time, or treatment group.

```{r newdata}
d <- readr::read_csv(here("./allmerged_posonly.csv"))
nrow(d)
d %<>% filter(!is.na(Death), !is.na(DSSTDY), !is.na(treatgroup))
nrow(d)
```

Time distribution: 

- remove one outlier, 
- among deaths, the treatment group seemed to have a longer survival time overall

```{r svtime}
f1 <- d %>%
    ggplot(aes(factor(Death), DSSTDY, fill = factor(treatgroup))) +
    geom_boxplot()

d %<>% filter(DSSTDY < 100)

f2 <- d %>%
    ggplot(aes(factor(Death), DSSTDY, fill = factor(treatgroup))) +
    geom_boxplot() +
    labs(title = 'remove time > 200')
f1 + f2
```

Though, only in GIN, treatment group seemed to have a longer survival time.

```{r}
d %>%
    ggplot(aes(factor(Death), DSSTDY, fill = factor(treatgroup))) +
    geom_boxplot() +
    facet_wrap(~COUNTRY)
```

Current way of treatment grouping, some study ID dont have both treatment groups.

```{r}
d %>%
    ggplot(aes(factor(Death), DSSTDY, fill = factor(treatgroup))) +
    geom_boxplot(position = position_dodge2(, "single")) +
    facet_wrap(~STUDYID) +
    coord_flip()
```

No relationship with age, it seems.

```{r}
d %>%
    ggplot(aes(AGE, DSSTDY, color = factor(Death))) +
    geom_point() +
    facet_wrap(~STUDYID)
```

What is sex `U`? also seems sex is not a concern. Remove sex `U` for now.

```{r}
d %>%
    ggplot(aes(SEX, DSSTDY, color = factor(SEX))) +
    geom_boxplot(position = position_dodge2(, "single")) +
    facet_wrap(~STUDYID)
d %<>% filter(SEX != 'U')
```

## Simple survival models

A hierarchical frailty survival model taking into potential correaltion within
country and study center effect. In general:

$$Survival\ time \sim LogLogistic(\alpha, \lambda)$$
$$\lambda = \exp(intercept + treat + f_{age} + sex + i_i)$$

where $i_i \sim N(0, \tau_c)$ are random effect of between individuals and 
$f_{age}$ is a 1st-order random-walk effect allowing exploring nonlinear trends.

> positive coefficient means **reducing** median survial time
> negative coefficient means **increasing** median survial time

```{r inla}
# install.packages("INLA",repos=c(getOption("repos"),INLA="https://inla.r-inla-download.org/R/testing"), dep=TRUE)
library(INLA)

formula <- inla.surv(DSSTDY, Death) ~ 1 + treatgroup +
    f(AGE, model = 'iid') + SEX + f(STUDYID) + f(iid)

d_inla <- d[, c("STUDYID", "COUNTRY", "SEX", "AGE", "Death", "DSSTDY", "treatgroup")] %>%
    mutate(iid = 1:n())

r = inla(formula,
    family = "loglogisticsurv", data = d_inla,
    control.family = list(variant = 1)
)

r$summary.fixed %>%
    select(1:5) %>%
    knitr::kable(digits = 2)


r$summary.hyperpar %>%
    select(1:5) %>%
    knitr::kable(digits = 2)

r$summary.random$AGE %>%
    as_tibble() %>%
    ggplot(aes(ID, mean, ymin = `0.025quant`, ymax = `0.975quant`)) +
    geom_line() +
    geom_ribbon(alpha = .3) +
    labs(title = 'Nonlinear effect of age on survival time')
```

Remarks:

- treatment increase survival time
- male had shorter survival time
- teenager had longer survival time than other ages
- Teenageer seems to have shorter survival time, and higher with age increase or decreasing.
